Tabular Constraint Learning

### 2016-04-15 Paper/ECAI/tabular.tex
reworking the statement and formalization

\newcommand{\constraints}{\ensuremath{\mathcal{C}}\xspace}
\newcommand{\format}[1]{\textit{#1}\xspace}
\newcommand{\generategroups}{\format{generateAssignments}}
\newcommand{\extractgroups}{\format{extractGroups}}
\newcommand{\extracttables}{\format{extractTables}}
\newcommand{\learnconstraints}{\format{learnConstraints}}
\newcommand{\findassignment}{\format{findSolutions}}
\newcommand{\postprocess}{\format{pruneRedundant}}
\newcommand{\constrainttorder}{\format{generalityOrder}}
\newcommand{\template}{\format{Constraint template}}

@@

Let us now define what is a tabular solution for a \template in general. Let $C$ denote a \template, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.

@@

\begin{minipage}[c]{14em}
  \vspace{5pt}
  \begin{tabular}{ll}
    \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
    \vspace{-4pt}
    &\\
    \textbf{Given:}& the set of all groups $\groups$ and of \template $\constraints$\\
    \textbf{Find:}&  all tabular solution $S$ in \groups for each constraint $c$ in \constraints \\
  \end{tabular}
  \vspace{6pt}
\end{minipage}

@@

  In theory each constraint is independent and equally useful. In practice, however, things are different. Assume that for some $g_x$ both \textit{series($g_x$)} and \textit{alldifferent($g_x$)} hold. From a user's perspective the last solution is useless, since he or she already knows that the series constraint implies the alldifferent constraint. To capture this condensed representation \cite{condensed} of solutions, we have introduced a specificity (generality) order between constraints, which is depicted using solid (dashed) arrows in Figure \ref{fig:learning_order}. This matters from implementation perspective as well, if we discovered that $g_x$ is not a permutation, then $g_x$ should not be tested for the series constraint at all. A solid lines from $C_1$ to $C_2$ indicates that $C_2$ is \textit{more specific} than $C_1$ and if $S$ is not a solution to $C_1$, then it is not a solution to $C_2$. A dashed arrow from $C_1$ to $C_2$ indicates that $C_1$ is \textit{a more specific} constraint than $C_2$ and therefore if $S$ is a solution for $C_1$ it should not be tested for $C_2$.

We call a tabular solution $S$ to a constraint $C$ \textit{the most specific} wrt \dependencies, iff there is no other constraint $C'$ for which $S$ is a solution and $C'$ is more specific in \dependencies.

@@

\begin{minipage}[c]{14em}
  \vspace{5pt}
  \begin{tabular}{ll}
    \multicolumn{2}{l}{{\textbf{Condensed Tabular Constraint Learning Problem}}}\\
    \vspace{-4pt}
    &\\
    \textbf{Given:}& the set of all groups $\groups$ and of \template's $\constraints$,\\
    & a constraint specificity DAG \dependencies \\
    \textbf{Find:}& all the most specific wrt \dependencies solutions $S$ in \groups\\
    & for each $c$ in \constraints \\
  \end{tabular}
  \vspace{6pt}
\end{minipage}

@@

\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- \template's, \dependencies -- dependency graph
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
    \State $S \gets \emptyset$ \Comment{The set of solutions}
  \For{$c \in \constrainttorder(\constraints,\dependencies)$}
  \State $v_1, ..., v_n =$ variables of $c$
    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S, \dependencies)$}
      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S, \dependencies)$
    \EndFor
  \EndFor\\
\Return $S$
\end{algorithmic}
\caption{Tabular constraint learning}
\label{algo:tcl}
\end{algorithm}
Essentially, our Algorithm \ref{algo:tcl} has three steps: constraint ordering, candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}. Let us elaborate on each step in detail.

@@

$\constrainttorder(\constraints,\dependencies)$ uses the DAG~\dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of~\dependencies.
Solid edges point from more general to more specific constraints.
Therefore, if there is a solid edge from constraint~$C_1$ to constraint~$C_2$, then a set of subgroups satisfying~$C_2$ must contain the subgroups satisfying~$C_1$.

Dashed edges represent the inverse relationship, pointing from specific to more general constraints.
If there is a dashed edge from $C_1$ to $C_2$, then a set of subgroups satisfying $C_2$ may not contain a subset that also satisfies $C_1$.

@@

Finding candidate groups for a \template~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
Several constraints (e.g. sum, max, count) have the same candidate generation procedure as they share the same \CSignature.

>>>>

\newcommand{\constraints}{\ensuremath{\mathcal{T}}\xspace}
\newcommand{\format}[1]{\textit{#1}\xspace}
\newcommand{\generategroups}{\format{generateAssignments}}
\newcommand{\extractgroups}{\format{extractGroups}}
\newcommand{\extracttables}{\format{extractTables}}
\newcommand{\learnconstraints}{\format{learnConstraints}}
\newcommand{\findassignment}{\format{findSolutions}}
\newcommand{\postprocess}{\format{pruneRedundant}}
\newcommand{\constrainttorder}{\format{generalityOrder}}
\newcommand{\template}{\format{Constraint template}}

@@

Let us introduce \template instantiation syntax. Let $S$ be a mapping from variables to the subgroups of \groups and $T$ be a \template, then $T^S$ is the template, in which all variables in all constrainst in \CSignature and \CFunction are instantiated with the corresponding subgroups from $S$. We refer to $T^S$ as a \textit{constraint} and to the instantiated \CSignature (\CFunction) as $\text{\CSignature}^S$ ($\text{\CFunction}^S$).

Let us now define what it means for a constraint from \template to hold in general. Let $T$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $T^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.

@@

\begin{minipage}[c]{14em}
  \vspace{5pt}
  \begin{tabular}{ll}
    \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
    \vspace{-4pt}
    &\\
    \textbf{Given:}& the set of all groups $\groups$ and of \template $\constraints$\\
    \textbf{Find:}&  all constraints $C$ over \groups for each template $t$ in \constraints \\
  \end{tabular}
  \vspace{6pt}
\end{minipage}

@@

  In theory each constraint is independent and equally useful. In practice, however, things are different. Assume that for some $g_x$ both \textit{series($g_x$)} and \textit{alldifferent($g_x$)} hold. From a user's perspective the last solution is useless, since he or she already knows that the series constraint implies the alldifferent constraint. To capture this condensed representation \cite{condensed} of solutions, we have introduced a specificity (generality) order between constraints, which is depicted using solid (dashed) arrows in Figure \ref{fig:learning_order}. This matters from implementation perspective as well, if we discovered that $g_x$ is not a permutation, then $g_x$ should not be tested for the series constraint at all. A solid lines from $t_1$ to $t_2$ indicates that $t_2$ is \textit{more specific} than $t_1$ and if $C_1=t_1^S$ from $t_1$ does not hold, then the constraint $C_2=t_2^S$ from $t_2$ does not hold neither. A dashed arrow from $t_1$ to $t_2$ indicates that $t_1$ is \textit{a more specific} constraint than $t_2$ and therefore if $C=t_1^S$ from $t_1$ holds it should not be tested for $t_2$.

Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, \dependencies be the DAG of constraint specificity, then the constraint $C$ equal to $t^S$ is called  \textit{the most specific} wrt \dependencies iff $C$ holds and there is no other template $t_{*}$ for which the constraint $C_{*}=t_{*}^S$ holds and $t_{*}$ is more specific in \dependencies.

@@

\begin{minipage}[c]{14em}
  \vspace{5pt}
  \begin{tabular}{ll}
    \multicolumn{2}{l}{{\textbf{Condensed Tabular Constraint Learning Problem}}}\\
    \vspace{-4pt}
    &\\
    \textbf{Given:}& the set of all groups $\groups$ and of \template's $\constraints$,\\
    & a constraint specificity DAG \dependencies \\
    \textbf{Find:}& all the most specific wrt \dependencies constraints $C$ over \groups\\
    & for each $t$ in \constraints \\
  \end{tabular}
  \vspace{6pt}
\end{minipage}

@@

\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- \template's, \dependencies -- dependency graph
    \State \textbf{Output:} $C$ -- learned constraints 
    \State $C \gets \emptyset$ \Comment{The set of constraints}
  \For{$t \in \constrainttorder(\constraints,\dependencies)$}
  \State $v_1, ..., v_n =$ variables of $T$
    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(t, \groups, C, \dependencies)$}
      \State $C \gets C \cup \findassignment(t, v_1{:}~G_1, \dots, v_n{:}~G_n, C, \dependencies)$
    \EndFor
  \EndFor\\
\Return $C$
\end{algorithmic}
\caption{Tabular constraint learning}
\label{algo:tcl}
\end{algorithm}
Essentially, our Algorithm \ref{algo:tcl} has three steps: constraint ordering, candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}. Let us elaborate on each step in detail.

@@

$\constrainttorder(\constraints,\dependencies)$ uses the DAG~\dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of~\dependencies.
Solid edges point from more general to more specific constraints.
Therefore, if there is a solid edge from constraint~$t_1$ to constraint~$t_2$, then a set of subgroups satisfying~$t_2$ must contain the subgroups satisfying~$t_1$.

Dashed edges represent the inverse relationship, pointing from specific to more general constraints.
If there is a dashed edge from $t_1$ to $t_2$, then a set of subgroups satisfying $t_2$ may not contain a subset that also satisfies $t_1$.

@@

Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $t$.
Several constraints (e.g. sum, max, count) have the same candidate generation procedure as they share the same \CSignature.
